---
title: "LTL Shipments EDA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(tidyverse)
library(modelr)
library(broom)
library(dplyr)
```

```{r}
library(xlsx)
CLient_Data <- read.xlsx("LTL_Dataset.xlsx", 1, header = TRUE)
```


```{r}
glimpse(CLient_Data)

```

Adding the zeros to the front of the zip codes that did not transfer from excel:

```{r}
Client2 <- transform(CLient_Data, OrigZip = str_pad(as.character(CLient_Data$OrigZip), width = 5, side = "left", pad = "0"), DestZip = str_pad(as.character(CLient_Data$DestZip), width = 5, side = "left", pad = "0") )


```


```{r}
library(lubridate)

Client3 <- transform(Client2, TruckloadActivityDate = ymd(Client2$TruckloadActivityDate), EarliestActualDeliveryDate = ymd(Client2$EarliestActualDeliveryDate))

glimpse(Client3)
```

Adding a variable to represent cost per pound:
```{r}
Client4 <- Client3 %>% 
  mutate(CostPerPound = (FreightCharge/SummaryActualWeight))
```

initial plotting to reveal any potential outliers:

```{r}
plot(Client4$SummaryActualWeight, Client4$FreightCharge)
```
We see there is an outlier. Looking at the data, we see that there is an entry with the weight of 400,000. We imagine this was a data entry error and remove the outlier so it will not affect the analysis 

```{r}
Client4 <- filter(Client4, SummaryActualWeight != 400000)

ggplot(data = Client4, aes(x = Client4$SummaryActualWeight, y = Client4$FreightCharge)) + geom_point(colour = "darkblue") + geom_smooth(se=FALSE, colour = "black") + labs(title = "Weight vs. Freight Charge", x = "Weight(lbs)", y = "Freight Charge($)") + theme(axis.title = element_text(size = 12), plot.title = element_text(size = 15, face = "bold"))
```

Attempts at finding the lanes with highest/lowest cost per pound:

```{r}
Client4 %>%
  group_by(OrigZip,DestZip) %>%
  summarise(Avg_CostPerLb = round(mean(CostPerPound), 2)) %>%
  arrange(desc(Avg_CostPerLb))
```

From the above data, we can see that three of the lanes have unusual avg cost per pound values. Investigating further by using a filter function to analyze those specific lanes. 

```{r}
Potential_Outliers <- Client4 %>% select(OrigZip, DestZip, SummaryActualWeight, FreightCharge) %>% filter(OrigZip == "33064", DestZip == "34758")

head(Potential_Outliers)

Potential_Outliers2 <- Client4 %>% select(OrigZip, DestZip, SummaryActualWeight, FreightCharge) %>% filter(OrigZip == "77020", DestZip == "44515")

head(Potential_Outliers2)

Potential_Outliers3 <- Client4 %>% select(OrigZip, DestZip, SummaryActualWeight, FreightCharge) %>% filter(OrigZip == "77020", DestZip == "08085")

head(Potential_Outliers3)
```

We see that one of the SummaryActualWeight values in each lane is unusually low. 

For lane 33064, 34758 there is a zero for one of the SummaryActualWeight values
For lane 77020, 44515 there is only one entry and the SummaryActual weight value is a zero. 
    (As this lane only has one entry, we intend to remove that row without impact to the findings, however if it can be corrected we can add the entry back in)

For lane 77020, 08085 there are two entries. One of the entries has a SummaryActualWeight value of 44. I assume that this was a data entry error, probably meaning to be 44000, but will be removed until clarified. 

We remove each of the outliers that are found and rerun the original code. 

```{r}
Client5 <- Client4 %>%
  filter(SummaryActualWeight != 0, SummaryActualWeight != 44)

Lane <- Client5 %>% 
  group_by(OrigCity, OrigState, OrigZip,DestCity, DestState, DestZip) %>%
  summarise(Avg_CostPerLb = round(mean(CostPerPound), 2)) %>%
  arrange(desc(Avg_CostPerLb))

Lane
```
```{r}
library(kableExtra)

Lane <- head(Lane, 3)

Lane %>%
   kbl() %>%
  kable_classic_2("striped", full_width = F)
 
```



```{r}
Client5 %>% 
  group_by(OrigZip,DestZip) %>%
  summarise(Avg_CostPerLb = round(mean(CostPerPound), 2)) %>%
  ggplot(aes(x = Avg_CostPerLb)) + geom_histogram(binwidth = 0.01, color = "black", fill = "lightblue") + xlab("Average Cost/lb") + ylab("Number of Lanes")
                                            

```



```{r}
Lane2 <- Client5 %>% 
  group_by(OrigCity, OrigState, OrigZip,DestCity, DestState, DestZip) %>%
  summarise(Avg_CostPerLb = round(mean(CostPerPound),2)) %>%
  arrange(Avg_CostPerLb)

Lane2
```
```{r}
Lane2 <- head(Lane2, 12)

Lane2 %>%
  kbl() %>%
  kable_classic_2("striped", full_width = F)
  
```

Investigating shipping modes:


```{r}
Client5 %>%
  ungroup() %>%
  group_by(Mode) %>%
  summarise(Count = n()) %>%
  ungroup()
```
visual for LTL argument: 
 
```{r}
plot(Client5$SummaryActualWeight, Client5$CostPerPound )
```


Showing that using LTL for smaller shipments could potentially save money: 

```{r}
LTL <- Client5 %>%
  filter(SummaryActualWeight <= 10000)

LTL %>% 
  group_by(SummaryActualWeight, Mode) %>%
  summarise(AvgCostperLb = mean(FreightCharge/SummaryActualWeight)) %>%
  ggplot(aes(SummaryActualWeight, AvgCostperLb, color = Mode)) + geom_point() + labs(x = "Weight(lbs)", y = "Average Cost/lb") 
  
LTL %>% 
  ungroup() %>%
  ggplot(aes(SummaryActualWeight, CostPerPound, color = Mode)) + geom_point() + labs(x = "Weight(lbs)", y = "Cost/lb") 




```


```{r}
Client5 %>%
  group_by(SummaryActualWeight,Mode) %>%
  summarise(AvgCostperLb = mean(FreightCharge/SummaryActualWeight)) %>%
  arrange(SummaryActualWeight) %>% 
  ungroup()
```

```{r}
Client5 %>% 
  group_by(OrigZip,DestZip, Mode) %>%
  summarise(Avg_CostPerLb = round(mean(CostPerPound),2)) %>%
  arrange(Avg_CostPerLb) %>%
  ungroup()
```

Investigating accesorial charges:

```{r}
##accesorial investigation 

Accesorial <- Client4 %>% 
  filter(AccessorialOtherCharge > 0)

Acc <- Accesorial %>%
  group_by(OrigCity, OrigState, OrigZip, DestCity, DestState, DestZip) %>%
  summarise(ExtraCharges = sum(AccessorialOtherCharge)) %>%
  arrange(desc(ExtraCharges)) %>%
  ungroup()

Acc

total_Accessorials <- sum(Accesorial$AccessorialOtherCharge)

Acc <- head(Acc, 5)

top5 <- sum(Acc$ExtraCharges)
top5


Acc %>% 
  kbl() %>%
  kable_classic_2("striped", full_width = F)
  
```

```{r}
Accesorial %>% 
  filter(OrigZip == "06437", DestZip == "20723")
```

```{r}
Client4 %>%
  select(OrigZip, DestZip, AccessorialOtherCharge, X.405...Fuel.Surcharge,Accessorial) %>%
  filter(OrigZip == "77020", DestZip == "78219")
```

```{r}
Client5 %>%
  select(OrigZip, OrigState, DestZip, DestState, SummaryActualWeight, Mode, FreightCharge, CostPerPound) %>%
  filter(OrigZip == "90620" | DestZip == "90620")
```
```{r}
Total_Freight_Spend <- sum(Client3$FreightCharge)

Total_Freight_Spend

Total_Overall_Spend <- sum(Client3$TotalCharges)

Total_Overall_Spend

Overall_AvgCostperLb <- mean(Client5$CostPerPound)

Overall_AvgCostperLb
```

```{r}
Client5 %>%
  filter(OrigState == "CT" | OrigState == "NJ", DestState == "NJ" | DestState == "CT") %>%
   group_by(OrigZip, DestZip) %>%
  summarise(AvgCostperLb = mean(CostPerPound)) %>%
  ggplot(aes(AvgCostperLb)) + geom_histogram(binwidth = 0.05, color = "black", fill = "darkblue") + labs(title = "Average Cost/lb Frequencies for NJ <-> CT Lanes", x = "Average Cost/lb", y = "Number of Lanes") + theme_classic() + theme(axis.title = element_text(size = 12), plot.title = element_text(size = 15, face = "bold"))
  
```

